Towards open collaboration in biomedical and health research: An open collaborative scoping review

### 2024-10-22 doc/paper.qmd
docs: small note about using open alex instead
started to add text from protocol and make structure for introduction

## Aims

Our original overall aim, as defined in the protocol, of this scoping
review is to identify current practices of open collaboration in
biomedical and health research.

@@

@@

@@

-   **Original timeline**: We were quite ambitious with the timeline for
    completing this scoping review. We did not adhere to the timeline.
-   **API vs web search results**: Almost all source databases had
    exposed APIs, with R packages available to access them. We chose to
    use these rather than using the web to extract the source data is
    due to our secondary aim. However, we noticed that there were some
    small differences in search results between the API query compared
    to the web-based query. We can't explain these differences, though
    they were quite minor.
-   **Exclude Web of Science source**: We were unable to connect the Web
    of Science source to R through the `{wosr}` package. Using the API
    key did not as documented in the package did not work. We decided
    that in the interest of reproducibility, that we exclude this
    source.

>>>>

-   Open science and its role in research

-   Collaboration and its role in modern research

    -   collaboration spanning from lab, to research center to multiple
        centers
    -   collaboration needed to tackle limitations: small populations,
        lack of replication, etc
    -   re-inventing discovery: a new way of working in science

Scientific research now almost always requires working with other
people.

With the growing complexity and specialization in scientific practices
and methods, together with globalisation of health and environmental
issues, there is a great need for a paradigm shift in research
collaboration to be able to tackle these challenges.

-   Open collaboration - a combination of two themes

    -   definition
    -   big potential in science
    -   but there very few resources and examples of how to integrate
        open science into collaborations

We define open collaboration using the definition as found in Forte and
Lampe (2013):

> "*an online environment that (a) supports the collective production of
> an artifact (b) through a technologically mediated collaboration
> platform (c) that presents a low barrier to entry and exit and (d)
> supports the emergence of persistent but malleable social
> structures.*"

With the increasing emphasis on and demand for science to be more open,
how we collaborate together is a key component to making science more
open from the start of any project. But how do we collaborate in an open
and transparent way? What are the best practices and tools we can use?
What is an ideal collaborative workflow and how close or far are we from
this ideal in reality?

-   Aim

This scoping review will focus on current practices of open
collaboration and open science in relation to collaboration in the field
of biomedical and health research.

@@

This is a scoping review, so we are following the framework described by
Arksey and O'Malley
([DOI:10.1080/1364557032000119616](https://doi.org/10.1080/1364557032000119616))
and by Levac, Colquhoun, and O'Brien
([DOI:10.1186/1748-5908-5-69](https://doi.org/10.1186/1748-5908-5-69))
as well as the guidelines outlined in the PRISMA-ScR
([DOI:10.7326/M18-0850](https://doi.org/10.7326/M18-0850) statement.

@@

We developed the initial search strategy in consultation with a research
librarian. We will collect the data via systematic searches of
databases, as well as some hand searches.

We used OpenAlex is an open database with scholarly works from MAG,
Crossref, ORCID, ROR, DOAJ, Unpaywall, Pubmed, Pubmed Central, The ISSN
International Centre, Internet Archive, Web crawls, Subject-area and
institutional repositories from arXiv and Zenodo.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(stringr)
library(here)
source(here("R/search-terms.R"))
search_terms_inline <- str_replace_all(search_terms$general, "\\*", "\\\\\\*")
```

We used the following search terms:

```         
> `r search_terms_inline`
```

Inclusion criteria included any document where open collaboration
practices are not the primary focus. We will be guided by the definition
of open collaboration from Forte and Lampe (2013) mentioned above in
determining whether the records are relevant. Any published document
with reporting on current open collaboration practices. Any published
document with advice, guidance, tools, and/or recommendations for
improving open collaboration. Article language in English.

Exclusion criteria were documents that do not report on specific open
collaboration practices. Documents published \>5 years in the past
(defined as before xxxx) to capture current/most recent practices.

All data obtained from all searches in online databases will be exported
into text files or CVS files for further processing including removal of
duplicates.

At least two of us will extract data using a standardized and tested
template. Data regarding the data source (e.g., author, title,
publication year), open collaboration practices, and any other relevant
information, will be extracted. Extracted data will be summarized with
the descriptive analytical method described by Arksey and O'Malley
(2005). The descriptive analytical method is aimed at identifying and
summarizing different open collaboration practices.

@@

-   **Original timeline**: We were quite ambitious with the timeline for
    completing this scoping review. We did not adhere to the timeline.
-   **API vs web search results**: Almost all source databases had
    exposed APIs, with R packages available to access them. We chose to
    use these rather than using the web to extract the source data is
    due to our secondary aim. However, we noticed that there were some
    small differences in search results between the API query compared
    to the web-based query. We can't explain these differences, though
    they were quite minor.
-   **Exclude Web of Science source**: We were unable to connect the Web
    of Science source to R through the `{wosr}` package. Using the API
    key did not as documented in the package did not work. We decided
    that in the interest of reproducibility, that we exclude this
    source.
-   **Move all searching to using Open Alex**

## Discussion
